Spatial variation in allometric growth of invasive lionfish has management implications

### 2019-04-02 docs/Manuscript.Rmd
Updated all, ready to release

---
title: "Spatial variation in allometric growth of invasive lionfish has management implications"
author: "Juan Carlos Villaseñor-Derbez, Sean P. Fitzgerald"
institution: "Bren School of Environmental Sciences & Management, University of California Santa Barbara, Santa Barbara, California, U.S."
abstract: |
  Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth, but sedentary species like lionfish often exhibit high levels of spatial variation in life history characteristics. We reviewed 17 published length-weight relationships for lionfish taken throughout their invasive range and found regional differences that led to significant misestimates when calculating weight from length observations. The spatial pattern we observed is consistent with findings from other studies focused on genetics or length-at-age. Here, the use of \textit{ex situ} parameter values resulted in total biomass estimates between 76.2% and 140% of true observed biomass, and up to a threefold under- or overestimation of total weight for an individual organism. These findings can have implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.
keywords:
  - Lionfish
  - invasive species
  - length-weight
  - allometric growth
  - regional variations
bibliography: references.bib
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
---

@@

We obtained a total of 18 parameter pairs by combining length-weight parameters extracted from the literature and the additional pair calculated here (Fig \ref{fig:all_allo}). Recall that the objective of this study is not to describe differences in the length-weight relationship between populations (which would require access to raw data), but rather to assess how *ex situ* parameter values influence the accuracy of weight estimates for lionfish, using the central Mexican Caribbean as a case study. Using each of the 18 parameter pairs, we estimated $TW$ from the $TL$ observations collected in the central Mexican Caribbean (n = 109, with $TL$ ranging from 34 mm to 310 mm) and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios were tested with an analysis of covariance (ANCOVA):

@@

The length-weight relationship for organisms from the central Mexican Caribbean (Fig \ref{fig:l-w-carib}) resulted in coefficient values of $a = 3.205 \times 10^{-6}$ and $b = `r bs`$ ($R^2 = `r Rsq`$, $F_{1, 107} = 6928.67; p < 0.001$). The allometric factor ($b$) was significantly different from $b = 3$ ($t_{107} = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$), corroborating that lionfish present allometric growth. The length-weight coefficients estimated here were within the range identified by studies from other regions (Table \ref{tab:all_params}).

@@

ANCOVA results revealed significant differences in our predicted weight ratios for the central Mexican Caribbean when using each of the different pairs of parameters ($F_{17, 1943} = 24.96; p < 0.001$; Fig \ref{fig:bio_ratio}). For example, the actual observed weights of the 109 lionfish from the Central Mexican Caribbean had a mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Weight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Weight) %>% formatC(digits = 2, format = "f")` g. However, if we used allometric parameter values from Banco Chinchorro in the Caribbean to predict weights from our observed length observations, we estimated a mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{sabidoitz_2016}. If we similarly used parameter values from North Carolina in the Western Atlantic to estimate lionfish weights in the Central Mexican Caribbean, we found a mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{barbour_2011}. Weights predicted from these extreme parameters correspond to mean predicted-to-observed weight ratios of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD), respectively. Furthermore, largest errors for individual organisms collected in the central Mexican Caribbean resulted in ratios of `r min(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r max(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` (*i.e.,* the tails of each violin in Fig \ref{fig:bio_ratio}). If we examined biomass (*i.e.,* summing across all 109 organisms) instead of mean ratios, total biomass estimates were 76.2% (4,363.53 g) and 140% (8,039.96 g) of true observed biomass (5,729.34 g). Parameters for this study estimate total biomass at 98% of observed biomass. These misestimates come from the two most extreme sets of parameters, but results varied consistently across locations (Figs \ref{fig:bio_ratio} and \ref{fig:errors}). Overall, the use of *ex situ* parameters led to significantly erroneous estimates of individual weight and total biomass for lionfish.

@@

```{r fit2, fig.height = 3.8, fig.pos = "H", fig.cap = "\\label{fig:all_allo}Length-weight relationships (n = 18) for 12 studies and this study. The curves are shown for the range of lengths reported in each study (See Supplementary Table 2); when ranges were not present, we use the ones found in this study (34 mm - 310 mm). Colors indicate studies from which the parameters were extracted. Dotted, dashed, and solid lines show models for males, females, and combined sexes, respectively. Letters in parentheses indicate if the study comes from the Gulf of Mexico (GoM), Western Atlantic (WA), or Caribbean (Ca). The dashed black line represents the relationship estimated in this study. There are two solid green lines for Sabido-Itza et al, 2016b, one for each of the two sites for which they report parameters. A log-log version of this figure is presented in Figure S4."}
fig_all_allo
```

@@

@@

Conflict of Interest: The authors declare that they have no conflict of interest.

>>>>

---
title: "Spatial variation in allometric growth of invasive lionfish has management implications"
author: "Juan Carlos Villaseñor-Derbez, Sean P. Fitzgerald"
institution: "Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, United States"
abstract: |
  Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth, but sedentary species like lionfish often exhibit high levels of spatial variation in life history characteristics. We reviewed 17 published length-weight relationships for lionfish taken throughout their invasive range and found regional differences that led to significant misestimates when calculating weight from length observations. The spatial pattern we observed is consistent with findings from other studies focused on genetics or length-at-age. Here, the use of \textit{ex situ} parameter values resulted in total biomass estimates between 76.2% and 140% of true observed biomass, and up to a threefold under- or overestimation of total weight for an individual organism. These findings can have implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.
keywords:
  - Lionfish
  - invasive species
  - length-weight
  - allometric growth
  - regional variations
bibliography: references.bib
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
---

@@

We obtained a total of 18 parameter pairs by combining length-weight parameters extracted from the literature and the additional pair calculated here (Table \ref{tab:all_params}). Recall that the objective of this study is not to describe differences in the length-weight relationship between populations (which would require access to raw data), but rather to assess how *ex situ* parameter values influence the accuracy of weight estimates for lionfish, using the central Mexican Caribbean as a case study. Using each of the 18 parameter pairs, we estimated $TW$ from the $TL$ observations collected in the central Mexican Caribbean (n = 109, with $TL$ ranging from 34 mm to 310 mm) and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios were tested with an analysis of covariance (ANCOVA):

@@

The length-weight relationship for organisms from the central Mexican Caribbean (Fig \ref{fig:l-w-carib}) resulted in coefficient values of $a = 3.205 \times 10^{-6}$ and $b = `r bs`$ ($R^2 = `r Rsq`$, $F_{1, 107} = 6928.67; p < 0.001$). The allometric factor ($b$) was significantly different from $b = 3$ ($t_{107} = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$), corroborating that lionfish present allometric growth. The length-weight coefficients estimated here were within the range identified by studies from other regions (Fig {fig:all_allo}, Table \ref{tab:all_params}).

@@

ANCOVA results revealed significant differences in our predicted weight ratios for the central Mexican Caribbean when using each of the different pairs of parameters ($F_{17, 1943} = 24.96; p < 0.001$; Fig \ref{fig:bio_ratio}). For example, the actual observed weights of the 109 lionfish from the Central Mexican Caribbean had a mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Weight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Weight) %>% formatC(digits = 2, format = "f")` g. However, if we used allometric parameter values from Banco Chinchorro in the Caribbean to predict weights from our observed length observations, we estimated a mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{sabidoitz_2016}. If we similarly used parameter values from North Carolina in the Western Atlantic to estimate lionfish weights in the Central Mexican Caribbean, we found a mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{barbour_2011}. Weights predicted from these extreme parameters correspond to mean predicted-to-observed weight ratios of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD), respectively. Furthermore, largest errors for individual organisms collected in the central Mexican Caribbean resulted in ratios of `r min(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r max(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` (*i.e.,* the tails of each violin in Fig \ref{fig:bio_ratio}). If we examined biomass (*i.e.,* summing across all 109 organisms) instead of mean ratios, total biomass estimates were 76.2% (4,363.53 g) and 140% (8,039.96 g) of true observed biomass (5,729.34 g). Parameters for this study estimate total biomass at 98% of observed biomass (Fig \ref{fig:errors}). These misestimates come from the two most extreme sets of parameters, but results varied consistently across locations (Figs \ref{fig:bio_ratio} and \ref{fig:errors}). Overall, the use of *ex situ* parameters led to significantly erroneous estimates of individual weight and total biomass for lionfish.

@@

```{r fit2, fig.height = 3.8, fig.pos = "H", fig.cap = "\\label{fig:all_allo}Length-weight relationships (n = 18) for 12 studies and this study. The curves are shown for the range of lengths reported in each study (See Supplementary Table 2); when ranges were not present, we use the ones found in this study (34 mm - 310 mm). Colors indicate studies from which the parameters were extracted. Dotted, dashed, and solid lines show models for males, females, and combined sexes, respectively. Letters in parentheses indicate if the study comes from the Gulf of Mexico (GoM), Western Atlantic (WA), or Caribbean (Ca). The dashed black line represents the relationship estimated in this study. There are two solid green lines for Sabido-Itza et al, 2016b, one for each of the two sites for which they report parameters. A log-log version of this figure is presented in Figure S1."}
fig_all_allo
```

@@

# Conclusions

@@

### 2019-04-02 docs/Manuscript.tex
Updated all, ready to release

\author[1]{Juan Carlos Villaseñor-Derbez}
\author[1]{Sean Fitzgerald}
\affil[1]{Bren School of Environmental Sciences and Management, University of California
  Santa Barbara, Santa Barbara, California, USA}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{juancarlos@ucsb.edu}

@@

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the Western Atlantic Ocean and Caribbean Sea, likely introduced
through release of aquarium-kept organisms \citep{betancurr_2011}.
Lionfish are the first invasive marine vertebrates established along
these coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016}, and
they have established invasive populations in coral reefs, estuaries,
mangroves, hard-bottomed areas, and mesophotic reefs
\citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.
Their presence has been labeled as a ``major marine invasion'' because
they threaten local biodiversity, spread rapidly, and are difficult to
manage \citep{hixon_2016}.

A substantial amount of research describes lionfish impacts throughout
their invaded range. A meta-analysis by \citet{peake_2018} showed that
invasive lionfish prey on at least 167 different species across the
tropical and temperate Western Atlantic. Their feeding behavior and high
consumption rates can reduce recruitment and population sizes of native
reef-fish species, and can further endanger reef fish
\citep[][but see \citealt{hackerott_2017} for a counterexample]{green_2012,rocha_2015}.
For example, field experiments showed that lionfish establishment in the
Bahamas led to reduced recruitment of native fishes by nearly 80\% over
a five-week period \citep{albins_2008}, and prey fish biomass declined
by 65\% over two years as lionfish biomass increased along Bahamian
coral reefs \citep{green_2012}. However, trophic impacts of lionfish can
be minimized if their local biomass is controlled by culling
\citep{ariasgonzalez_2011}.

@@

We obtained a total of 18 parameter pairs by combining length-weight
parameters extracted from the literature and the additional pair
calculated here (Fig \ref{fig:all_allo}). Recall that the objective of
this study is not to describe differences in the length-weight
relationship between populations (which would require access to raw
data), but rather to assess how \emph{ex situ} parameter values
influence the accuracy of weight estimates for lionfish, using the
central Mexican Caribbean as a case study. Using each of the 18
parameter pairs, we estimated \(TW\) from the \(TL\) observations
collected in the central Mexican Caribbean (n = 109, with \(TL\) ranging
from 34 mm to 310 mm) and divided predicted weights by known observed
weights to obtain a simple measure of over- or underestimation.
Difference in mean weight ratios were tested with an analysis of
covariance (ANCOVA):

@@

The length-weight relationship for organisms from the central Mexican
Caribbean (Fig \ref{fig:l-w-carib}) resulted in coefficient values of
\(a = 3.205 \times 10^{-6}\) and \(b = 3.235\) (\(R^2 = 0.977\),
\(F_{1, 107} = 6928.67; p < 0.001\)). The allometric factor (\(b\)) was
significantly different from \(b = 3\) (\(t_{107} = 6.04; p<0.001\)),
corroborating that lionfish present allometric growth. The length-weight
coefficients estimated here were within the range identified by studies
from other regions (Table \ref{tab:all_params}).

@@

ANCOVA results revealed significant differences in our predicted weight
ratios for the central Mexican Caribbean when using each of the
different pairs of parameters (\(F_{17, 1943} = 24.96; p < 0.001\); Fig
\ref{fig:bio_ratio}). For example, the actual observed weights of the
109 lionfish from the Central Mexican Caribbean had a mean \(\pm\) SD of
52.56 \(\pm\) 76.58 g. However, if we used allometric parameter values
from Banco Chinchorro in the Caribbean to predict weights from our
observed length observations, we estimated a mean \(\pm\) SD of 40.37
\(\pm\) 58.74 g \citep{sabidoitz_2016}. If we similarly used parameter
values from North Carolina in the Western Atlantic to estimate lionfish
weights in the Central Mexican Caribbean, we found a mean \(\pm\) SD of
73.76 \(\pm\) 96.11 g \citep{barbour_2011}. Weights predicted from these
extreme parameters correspond to mean predicted-to-observed weight
ratios of 0.80 \(\pm\) 0.19 and 1.76 \(\pm\) 0.50 (mean \(\pm\) SD),
respectively. Furthermore, largest errors for individual organisms
collected in the central Mexican Caribbean resulted in ratios of 0.36
and 3.51 (\emph{i.e.,} the tails of each violin in Fig
\ref{fig:bio_ratio}). If we examined biomass (\emph{i.e.,} summing
across all 109 organisms) instead of mean ratios, total biomass
estimates were 76.2\% (4,363.53 g) and 140\% (8,039.96 g) of true
observed biomass (5,729.34 g). Parameters for this study estimate total
biomass at 98\% of observed biomass. These misestimates come from the
two most extreme sets of parameters, but results varied consistently
across locations (Figs \ref{fig:bio_ratio} and \ref{fig:errors}).
Overall, the use of \emph{ex situ} parameters led to significantly
erroneous estimates of individual weight and total biomass for lionfish.

Tukey's \emph{post-hoc} test showed that weight ratios for the central
Mexican Caribbean differeed from those obtained with parameters from the
Western Atlantic, and most sites in the Caribbean and the Gulf of Mexico
(Tukey's HSD \(p > 0.05\)). The only sites where weight ratios did not
differ from the central Mexican Caribbean were Little Cayman
\citep{edwards_2014}, Bahamas \citep{darling_2011},
and the Northern Gulf of Mexico (\citet{dahl_2014}; Tukey's HSD
\(p > 0.05\)). All weight estimates using parameters from the Gulf of
Mexico and Western Atlantic were higher than observed values, and only
parameters from the Caribbean produced weights smaller than observed
(Fig \ref{fig:bio_ratio}). The regional average (\(\pm\) SD) of
predicted-to-observed weight ratios from these three regions were 1.24
\(\pm\) 0.309, 1.41 \(\pm\) 0.523, and 1.20 \(\pm\) 0.423 for the Gulf
of Mexico, Western Atlantic, and Caribbean, respectively. This suggests
that the smallest errors are observed when using parameters from other
locations in the Caribbean.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/fit2-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 18) for 12
studies and this study. The curves are shown for the range of lengths
reported in each study (See Supplementary Table 2); when ranges were not
present, we use the ones found in this study (34 mm - 310 mm). Colors
indicate studies from which the parameters were extracted. Dotted,
dashed, and solid lines show models for males, females, and combined
sexes, respectively. Letters in parentheses indicate if the study comes
from the Gulf of Mexico (GoM), Western Atlantic (WA), or Caribbean (Ca).
The dashed black line represents the relationship estimated in this
study. There are two solid green lines for Sabido-Itza et al, 2016b, one
for each of the two sites for which they report parameters. A log-log
version of this figure is presented in Figure S4.}
\end{figure}

@@

Our results suggest that lionfish exhibit highly variable, spatially
heterogeneous allometric relationships across their invaded range, and
that this variation is relevant for managing invasions. Moreover, we
show that the use of \emph{ex situ} parameter values may lead to highly
biased weight and total biomass estimates. Our comparison of observed
weights to those predicted with locally-informed parameters and \emph{ex
situ} parameters showed that weight of an individual lionfish can be
overestimated by more than threefold, highlighting the need to use local
information. Here we discuss the implications of our findings, possible
shortcommings in our analyses, and highlight potential future research
directions.

@@

We detected substantial differences in weight-at-length between
organisms from the Caribbean, Gulf of Mexico, and Western Atlantic.
Groupings of predicted-to-observed weight ratios identified in our
\emph{post hoc} testing aligned with the spatial distribution of the
examined studies, suggesting that these differences may be mediated by
space. These regional allometric differences mirror similar patterns in
length-at-age of lionfish across both their invaded and native regions
\citep{pusack_2016}. Variation may be driven by genetics or by
organisms' exposure to distinct environmental conditions. For example,
\citet{betancurr_2011} used mitochondrial DNA to demonstrate the
existence of two distinct population groups, identified as the
``Caribbean group'' and ``Northern Group'', and \citet{fogg_2015}
alternatively suggested that length-at-age differences may be driven by
the environment.

One might be inclined to attribute all variation in the lionfish
length-weight relationship to the spatial origin of these parameters.
However, samples from the 12 studies included here were not only
collected in different locations, but also at different points in time
and across different depth and size ranges (See Supplementary Table 2
for an extended version of Table 1). The magnitude of the bias
discovered in this study and our lack of understanding the sources
driving spatial variation for lionfish highlights the need to
simultaneously collect length-weight information across the invaded
range to test for spatially-induced patterns and to link these findings
to previously suggested environmental and genetic structures. Such an
endeavor would provide insight into lionfish biology and better inform
management. However, while we could not evaluate how these factors
influenced length-weight estimates from previous studies without raw
data, we still show that a lack of locally-calculated parameters can
induce significant bias when calculating weight from length
observations. We demonstrate the importance of using \emph{in situ}
parameters to obtain accurate weight estimates regardless of the
underlying mechanisms driving variation between populations.

@@

The results presented here have key implications for management. For
example, \citet{edwards_2014} simulated a lionfish culling program under
two scenarios, one using length-at-age and length-to-weight parameters
from North Carolina and one using parameters from Little Cayman. Their
results show that using different parameters caused up to a four-year
difference in the time required for the simulated lionfish population to
recover to 90\% of its initial biomass after removals ceased. Here, we
show that using one set of length-weight parameters versus another for a
given length can result in more than a threefold under- or
overestimation of total weight for individual fish, and that total
biomass estimates may range between 76\% and 140\% of true observed
biomass. These differences become especially important when allocating
resources for lionfish removal programs, incentivizing lionfish
fisheries as a source of alternative livelihoods, or estimating
ecosystem impacts. Research efforts focused on invasive lionfish
populations need to use parameters calculated for their region to the
extent possible, or use different sets of parameters that provide
appropriate upper and lower bounds in their results.

>>>>

\author[1]{Juan Carlos Villaseñor-Derbez}
\author[1]{Sean Fitzgerald}
\affil[1]{Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, United States}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{juancarlos@ucsb.edu}

@@

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the Western Atlantic Ocean and Caribbean Sea, likely introduced
through release of aquarium-kept organisms \citep{betancurr_2011}.
Lionfish are the first invasive marine vertebrates established along
these coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016}, and
they have established populations in coral reefs, estuaries,
mangroves, hard-bottomed areas, and mesophotic reefs
\citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.
Their presence has been labeled as a ``major marine invasion'' because
they threaten local biodiversity, spread rapidly, and are difficult to
manage \citep{hixon_2016}.

A substantial amount of research describes lionfish impacts throughout
their invaded range. A meta-analysis by \citet{peake_2018} showed that
invasive lionfish prey on at least 167 different species across the
tropical and temperate Western Atlantic. Their feeding behavior and high
consumption rates can reduce recruitment and population sizes of native
reef-fish species, and can further endanger reef fish
\citep[][but see \citealt{hackerott_2017} for a counterexample]{green_2012,rocha_2015}.
For example, field experiments showed that lionfish in the
Bahamas led to reduced recruitment of native fishes by nearly 80\% over
a five-week period \citep{albins_2008}, and prey fish biomass declined
by 65\% over two years as lionfish biomass increased along Bahamian
coral reefs \citep{green_2012}. However, trophic impacts of lionfish can
be minimized if their local biomass is controlled by culling
\citep{ariasgonzalez_2011}.

@@

We obtained a total of 18 parameter pairs by combining length-weight
parameters extracted from the literature and the additional pair
calculated here (Table \ref{tab:all_params}). Recall that the objective of
this study is not to describe differences in the length-weight
relationship between populations (which would require access to raw
data), but rather to assess how \emph{ex situ} parameter values
influence the accuracy of weight estimates for lionfish, using the
central Mexican Caribbean as a case study. Using each of the 18
parameter pairs, we estimated \(TW\) from the \(TL\) observations
collected in the central Mexican Caribbean (n = 109, with \(TL\) ranging
from 34 mm to 310 mm) and divided predicted weights by known observed
weights to obtain a simple measure of over- or underestimation.
Difference in mean weight ratios were tested with an analysis of
covariance (ANCOVA):

@@

The length-weight relationship for organisms from the central Mexican
Caribbean (Fig \ref{fig:l-w-carib}) resulted in coefficient values of
\(a = 3.205 \times 10^{-6}\) and \(b = 3.235\) (\(R^2 = 0.977\),
\(F_{1, 107} = 6928.67; p < 0.001\)). The allometric factor (\(b\)) was
significantly different from \(b = 3\) (\(t_{107} = 6.04; p<0.001\)),
corroborating that lionfish present allometric growth. The length-weight
coefficients estimated here were within the range identified by studies
from other regions (Fig \ref{fig:all_allo}, Table \ref{tab:all_params}).

@@

ANCOVA results revealed significant differences in our predicted weight
ratios for the central Mexican Caribbean when using each of the
different pairs of parameters (\(F_{17, 1943} = 24.96; p < 0.001\); Fig
\ref{fig:bio_ratio}). For example, the actual observed weights of the
109 lionfish from the Central Mexican Caribbean had a mean \(\pm\) SD of
52.56 \(\pm\) 76.58 g. However, if we used allometric parameter values
from Banco Chinchorro in the Caribbean to predict weights from our
observed length observations, we estimated a mean \(\pm\) SD of 40.37
\(\pm\) 58.74 g \citep{sabidoitz_2016}. If we similarly used parameter
values from North Carolina in the Western Atlantic to estimate lionfish
weights in the Central Mexican Caribbean, we found a mean \(\pm\) SD of
73.76 \(\pm\) 96.11 g \citep{barbour_2011}. Weights predicted from these
extreme parameters correspond to mean predicted-to-observed weight
ratios of 0.80 \(\pm\) 0.19 and 1.76 \(\pm\) 0.50 (mean \(\pm\) SD),
respectively. Furthermore, largest errors for individual organisms
collected in the central Mexican Caribbean resulted in ratios of 0.36
and 3.51 (\emph{i.e.,} the tails of each violin in Fig
\ref{fig:bio_ratio}). If we examined biomass (\emph{i.e.,} summing
across all 109 organisms) instead of mean ratios, total biomass
estimates were 76.2\% (4,363.53 g) and 140\% (8,039.96 g) of true
observed biomass (5,729.34 g). Parameters for this study estimate total
biomass at 98\% of observed biomass (Fig \ref{fig:errors}). These misestimates come from the
two most extreme sets of parameters, but results varied consistently
across locations (Figs \ref{fig:bio_ratio} and \ref{fig:errors}).
Overall, the use of \emph{ex situ} parameters led to significantly
erroneous estimates of individual weight and total biomass for lionfish.

Tukey's \emph{post-hoc} test showed that weight ratios for the central
Mexican Caribbean differed from those obtained with parameters from the
Western Atlantic in North Carolina \citep{barbour_2011},
and most sites in the Caribbean and the Gulf of Mexico
(Tukey's HSD \(p > 0.05\)). The only sites where weight ratios did not
differ from the central Mexican Caribbean were Little Cayman
\citep{edwards_2014}, Bahamas \citep{darling_2011},
and the Northern Gulf of Mexico (\citet{dahl_2014}; Tukey's HSD
\(p > 0.05\)). All weight estimates using parameters from the Gulf of
Mexico and Western Atlantic were higher than observed values, and only
parameters from the Caribbean produced weights smaller than observed
(Fig \ref{fig:bio_ratio}). The regional average (\(\pm\) SD) of
predicted-to-observed weight ratios from these three regions were 1.24
\(\pm\) 0.309, 1.41 \(\pm\) 0.523, and 1.20 \(\pm\) 0.423 for the Gulf
of Mexico, Western Atlantic, and Caribbean, respectively. This suggests
that the smallest errors are observed when using parameters from other
locations in the Caribbean.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/fit2-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 18) for 12
studies and this study. The curves are shown for the range of lengths
reported in each study (See Supplementary Table 2); when ranges were not
present, we use the ones found in this study (34 mm - 310 mm). Colors
indicate studies from which the parameters were extracted. Dotted,
dashed, and solid lines show models for males, females, and combined
sexes, respectively. Letters in parentheses indicate if the study comes
from the Gulf of Mexico (GoM), Western Atlantic (WA), or Caribbean (Ca).
The dashed black line represents the relationship estimated in this
study. There are two solid green lines for Sabido-Itza et al, 2016b, one
for each of the two sites for which they report parameters. A log-log
version of this figure is presented in Figure S1.}
\end{figure}

@@

Our results suggest that lionfish exhibit highly variable, spatially
heterogeneous allometric relationships across their invaded range, and
that this variation may be relevant for managing invasions. Moreover, we
show that the use of \emph{ex situ} parameter values may lead to highly
biased weight and total biomass estimates. Our comparison of observed
weights to those predicted with locally-informed parameters and \emph{ex
situ} parameters showed that weight of an individual lionfish can be
overestimated by more than threefold, highlighting the need to use local
information. Here we discuss the implications of our findings, possible
shortcommings in our analyses, and highlight potential future research
directions.

@@

We detected substantial differences in weight-at-length between
organisms from the Caribbean, Gulf of Mexico, and Western Atlantic.
Groupings of predicted-to-observed weight ratios identified in our
\emph{post hoc} testing aligned with the spatial distribution of the
examined studies, suggesting that these differences may be mediated by
location. These regional allometric differences mirror similar patterns in
length-at-age of lionfish across both their invaded and native regions
\citep{pusack_2016}. Variation may be driven by genetics or by
organisms' exposure to distinct environmental conditions. For example,
\citet{betancurr_2011} used mitochondrial DNA to demonstrate the
existence of two distinct population groups, identified as the
``Caribbean group'' and ``Northern Group'', and \citet{fogg_2015}
alternatively suggested that length-at-age differences may be driven by
the environment.

One might be inclined to attribute all variation in the lionfish
length-weight relationship to the spatial origin of these parameters.
However, samples from the 12 studies included here were not only
collected in different locations, but also at different times of the year
and across different depth and size ranges (See Supplementary Table 2
for an extended version of Table 1). The magnitude of the bias
discovered in this study and our lack of understanding the sources
driving spatial variation for lionfish highlights the need to
simultaneously collect length-weight information across the invaded
range to test for spatially-induced patterns and to link these findings
to previously suggested environmental and genetic structures. Such an
endeavor would provide insight into lionfish biology and better inform
management. However, while we could not evaluate how these factors
influenced length-weight estimates from previous studies without raw
data, we still show that a lack of locally-calculated parameters can
induce significant bias when calculating weight from length
observations. We demonstrate the importance of using \emph{in situ}
parameters to obtain accurate weight estimates regardless of the
underlying mechanisms driving variation between populations.

@@

\section*{Conclusion}

The results presented here have key implications for management. For
example, \citet{edwards_2014} simulated a lionfish culling program under
two scenarios, one using length-at-age and length-to-weight parameters
from North Carolina and one using parameters from Little Cayman. Their
results show that using different parameters caused up to a four-year
difference in the time required for the simulated lionfish population to
recover to 90\% of its initial biomass after removals ceased.

Here, we show that using one set of length-weight parameters versus another for a
given length can result in more than a threefold under- or
overestimation of total weight for individual fish, and that total
biomass estimates may range between 76\% and 140\% of true observed
biomass. These differences become especially important when allocating
resources for lionfish removal programs, incentivizing lionfish
fisheries as a source of alternative livelihoods, or estimating
ecosystem impacts. Research efforts focused on invasive lionfish
populations need to use parameters calculated for their region to the
extent possible, or use different sets of parameters that provide
appropriate upper and lower bounds in their results.
